An efficient contact value approach for the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-06-27 papers/pair-correlation/paper.tex
start adding new fit equation to pair paper Tested on bennet. Took 0:20.43.

\input{figs/fit-parameters}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-g2}
  \caption{Plot of the hard-sphere radial distribution function, with
    our separable fit.  The solid lines show the true radial
    distribution function $g(r)$, as found from Monte Carlo
    simulation, and the dashed lines show our separable fit to the
    radial distribution functions
    $g_S(r)$.}\label{fig:radial-distribution}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/short-range-ghs-alt}
  \caption{Plot of the hard-sphere radial distribution function, with
    our short-range separable
    fit.}\label{fig:short-range-radial-distribution}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-alt}
  \caption{Plot of the hard-sphere radial distribution function. The blue
    lines show our separable fit, the black dots are Monte Carlo data, and
    the dashed lines are results of the Gil-Villegas\cite{gil1997statistical}
    fit.\pahosays{An idea for an alternate version of the radial distribution
      figure.}}\label{fig:radial-distribution-alt}
\end{figure}

@@

%% We construct our function's form in terms of an expansion of powers of
%% $h_{\sigma} = g_{\sigma}-1$, ensuring that it approaches $g_{\sigma}$
%% in the contact distance limit, leaving our function with tunable
%% constant perameters.  We run a simple monte-carlo simulation to
%% calculate values of the homogeneous radial distribution function at
%% varying distance perameter $r$.  We then tune our function's constant
%% parameters by matching our function's results against the monte-carlo
%% data, using a least squares fit method.  During this process it is
%% simplest to use an analytic expression for $g_{\sigma}$ in our
%% function and so we use the well known and very successful Carnahan
%% Starling equation to calculate $g_{\sigma}$.  This is appropriate
%% since our published $g_{\sigma}$ approaches the Carnahan Starling
%% version in the homogeneous limit and it's this limit, along with the
%% contact limit, that we would like our constructed radial distribution
%% function to get right.  After this process we have an appropriate
%% radial distribution function that is dependent only on $r$ and
%% $g_{\sigma}$.

%% The function has the form:
%% \begin{multline}
%%   g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
%%   + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
%%   + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
%% \end{multline}
%% where $\kappaO$,$\kappaI$, and $\kappaZ$ are fitted constants and
%% $I(g_\sigma)$ is a function of $g_\sigma$.

%% The finalized form of our pair correlation function is:
%% \begin{align}
%%   g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
%% \end{align}

We construct $g_S(r, g_\sigma)$ to satisfy several constraints.
Naturally, we match the value at contact:
\begin{equation}
  g_S(\sigma, g_\sigma) = g_\sigma.
\end{equation}
A second constraint applies to the integral of the total correlation
function, $h(r) = g(r) - 1$:
\begin{align}
  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
\end{align}
where $\chi_T$ is the isothermal compressibility of the hard-sphere fluid.
We have found that the slope of the radial distribution function at
contact is given to a very good approximation by:
\begin{align}
  g'(\sigma) \approx g_\sigma (1 - g_\sigma)
\end{align}
and we apply this as a final constraint.\par
Under these three constraints, we applied a least-squares fit to
simulation data for the radial distribution function at packing
fractions from 0.05 to 0.4 in steps of 0.05.  Our fitted radial
distribution function is given by
\begin{multline}
  g_S(r,g_\sigma) = 1 + (g_\sigma-1) e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
  + (g_\sigma -1)(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
  + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
\end{multline}
\begin{equation}
  I(g_\sigma) = \kappaZ^5 \frac{
    \frac{\chi-1}{24\eta} + (1-g_\sigma) \frac{(\kappaO-2
      g_\sigma)(\kappaI^2 + 4 \kappaI + 6)}{\kappaI^4}
    + (1-g_\sigma)\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
  }{2 \kappaZ^2 + 12 \kappaZ + 24}
\end{equation}
\begin{equation}
  \chi = \frac{(1-\eta)^4}{\eta^4 - 4\eta^3 + 4\eta^2 + 4\eta + 1}
\end{equation}
where the dimensionless isothermal compressibility $\chi$ is obtained
from the Carnahan-Starling equation of state.
Finally, we can convert between the packing fraction $\eta$ and the
radial distribution function at contact $g_\sigma$ using the
Carnahan-Starling approximation for $g_\sigma$:
\newcommand\nastyetacuberoot{\sqrt[3]{54 g_\sigma^2 -
    6\sqrt{81g_\sigma^4 - 6g_\sigma^3}}}
\begin{align}
  g_\sigma &= \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}\text{, and its inverse} \\
  \eta &= 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}.
\end{align}
This function includes three fitted parameters, $\kappaO =
\kappazero$, $\kappaI = \kappaone$,
and $\kappaZ = \kappatwo$.
The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
figure, it is apparent that the constraint on the integral of $h(r)$ from
Eq.~\ref{eq:total-constraint} results in canceling errors at large
distances.  The maximum error in $g_S(r)$ at packing fractions up to
0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
\rmaxerr R$.

>>>>

\input{figs/short-range-fit-parameters}

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/short-range-ghs-alt}
  \caption{Plot of the hard-sphere radial distribution function. The
    blue lines show our separable fit, the black dots show the true
    radial distribution function $g(r)$ as found from Monte Carlo
    simulation, and the dashed lines are results of the
    Gil-Villegas\cite{gil1997statistical} fit.  The dotted extension
    of each fitted curve indicates the value of the function outside
    of the fitted region.  }\label{fig:radial-distribution}
\end{figure}

@@

We construct $g_S(r, g_\sigma)$ to satisfy several constraints.
Naturally, we match the value at contact:
\begin{equation}
  g_S(\sigma, g_\sigma) = g_\sigma.
\end{equation}

\begin{widetext}
\begin{align}
  g_S(r,g_\sigma) &=
  1 + \left(h_\sigma + \sum_{i=2}^{4} \sum_{j=1}^{4} h_\sigma^i
  \zeta^j\right)e^{-\alpha \zeta}
  \\
  &= \kappamatrix
\end{align}
\end{widetext}
where $h_\sigma = g_\sigma - 1$ and $\zeta \equiv
\frac{r}{\sigma}-1$.  \fixme{Why start at $i=2$ rather than $i=1$ for
  the fitted terms? This seems suboptimal.}

\fixme{Change code to round to a fixed (small) number of digits in fit.}
